Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-09-21 packages/spec-haskell/yellowpaper/Paper.tex
[yellowpaper] updated outline and added payment system section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document Preamble
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[a4paper]{article}

@@

%% Import Packages
\usepackage{fvextra}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{amsmath}
\usepackage{enumitem}

@@

\section{Introduction}

It should be fair to say, every aspect of money are controversial: the nature of money, the value of money, money and
banking, and monetary reconstruction. Two major schools of thoughts about theory of money are the \textit{Austrian
    school} \cite{von2013theory} and the \textit{Chicago school} \cite{friedman1989quantity}. That is before the
appearance of Internet-era version of monetary reconstruction, broadly defined as cryptocurrency, which challenges
theories of money further and demands their updates \cite{ammous2018can} \cite{hardle2020understanding}.

@@

\begin{quotation}
The function of money is to facilitate the business of the market by acting as a common medium of
exchange. \footfullcite[][Part One, Chapter I, § 1, p1]{von2013theory}
\end{quotation}

@@

In Part I, we shall first explore the foundation for the survey. Here we present a formal definitions of payment system,
and various models in the system. We then select a few relevant approaches used in computer science useful for defining
formal specification of payment system.

One of the approaches is denotational semantics, which is used in Part II of the paper to define the general payment
primitives. Along with the denotational semantics, a restatement of it in \textit{Haskell programming language} is also
included.

In Part III, a reference implementation of the general payment primitives and its payment system called
\textit{Superfluid Money} is introduced.

In Part IV, some possible further investigations are included for future study purpose.

\newpage
\part{Foundation}
\newpage

\section{Nature of Money}

Nature of money.

\section{Money Distribution}

@@

TODO...

\subsection{A New Unifying Theory}

@@

\section{Forms of Medium}

\section{Payment Models}

\subsection{Discrete Payments}

\subsection{Continuous Payments}

In contract, under the contextual money schemes, the liquidity value and maybe even bearer of a money unit can be a
function of context too.

TODO...

\section{Related Paradigms}

@@

\section{Generalized Denotational Payment Semantics}

@@

\newpage
\part{Superfluid Money}
\newpage

@@

\newpage

\printbibliography{}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Document Preamble
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[a4paper]{report}

@@

%% Import Packages
\usepackage{fvextra}
\usepackage{csquotes}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{graphicx}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

It should be fair to say, every aspect of money are controversial: the nature of money, the value of money, money and
banking, and monetary reconstruction. Two major schools of thoughts about theory of money are the \textit{Austrian
    school} (\cite{von2013theory}) and the \textit{Chicago school} (\cite{friedman1989quantity}). That is before the
appearance of Internet-era version of monetary reconstruction, broadly defined as cryptocurrency, which challenges
theories of money further and demands their updates (\cite{ammous2018can} \cite{hardle2020understanding}).

@@

\begin{quotation}
The function of money is to facilitate the business of the market by acting as a common medium of
exchange. \footfullcite[][Chapter One, Chapter I, § 1, p1]{von2013theory}
\end{quotation}

@@

In Chapter I, we shall first explore the foundation for the survey. Here we present a formal definition of payment system,
and various models used in the system. We then select a few relevant approaches used in computer science useful for
modeling and defining formal specification for the payment system.

One of the approaches is \textit{denotational semantics}, which is used in Chapter II of the paper to define the general
payment primitives. Along with the denotational semantics, a restatement\footnote{It is a borrowed term from common law:
"restatement of the law". In our case, the denotative mathematical "laws".} of it in \textit{Haskell programming
    language} (\cite{hudak1992report} \cite{jones2003haskell} \cite{marlow2010haskell}) is also included.

In Chapter III, a reference implementation of the general payment primitives and its payment system called
\textit{Superfluid Money} is introduced.

In Chapter IV, some possible further investigations are included for future study purpose.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Foundation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Payment System}

We shall define a payment system, whereby:

\begin{itemize}
\item \textit{money distribution} models how monetary value is distributed amongst bearers\footnote{(Banking \& Finance)
a person who presents a note or bill for payment. - Collins English Dictionary},
\item \textit{payment primitives} updates money distribution,
\item \textit{financial system} provides the execution environment for the payment primitives,
\item and \textit{forms of money medium} are the "user interfaces" of money for the bearers.
\end{itemize}

\subsection{Money Distribution Models}

@@

\begin{figure}
  \includegraphics[width=\linewidth]{assets/discrete-money-distribution.png}
  \caption{Schematic representation of discrete money distribution}
  \label{fig:discrete-md}
\end{figure}

\subsubsection{A New Unifying Model}

@@

\subsection{Payment Primitives}

\subsection{Payment System Models}

\subsubsection{Discrete Payments}

\subsubsection{Continuous Payments}

\subsection{Money Medium Models}

\section{Relevant Approaches}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{General Payment Primitives}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Denotational Semantics}

@@

\section{Restatement in Haskell}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Superfluid Money - A Reference Implementation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Future Investigations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Restatement in Agda and Correctness Proof}

\section{Constraint in Consensus}

\section{Payment in General Accounting Domains}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\printbibliography{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
